---
title: "Tutorial for custom geom to draw earthquake timeline"
author: "Coursera Learner 42"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo=FALSE}
library(CourseraRcapstone)
library(magrittr)
library(ggplot2)
library(grid)
options(width = 120)
```
## Read the clean data

Read the raw data manually then pass to the `eq_clean_data` function
```{r, message=FALSE, WARNING=FALSE}
raw_data_fn <- file.path(system.file("extdata", package="CourseraRcapstone"),"signif.txt")
data <- readr::read_delim(raw_data_fn, "\t") %>% eq_clean_data
str(data)
```

Or use directy the built in `read_eq_clean_data` function
```{r, message=FALSE, WARNING=FALSE}
data2 <- read_eq_clean_data(raw_data_fn)
identical(data, data2)
```

## Draw earthquake timeline

The follong snippet draws earthquakes observed in the US and Canada from 2000
```{r, fig.show='hold', message=FALSE, warning=FALSE}
    read_eq_clean_data() %>%
        dplyr::filter(YEAR > 2000 & !IS_BC & COUNTRY %in% c("USA", "CANADA")) %>%
        dplyr::mutate(TOTAL_DEATHS = as.numeric(TOTAL_DEATHS),
                      EQ_PRIMARY = as.numeric(EQ_PRIMARY)) %>%
        ggplot2::ggplot(ggplot2::aes(x = DATE,
                                     y = COUNTRY,
                                     colour = TOTAL_DEATHS,
                                     size = EQ_PRIMARY
                                     )) +
        geom_timeline() +
	theme_timeline() +
	labs(size = "Richter scale value", color = "# deaths", fill = "# deaths")
```

## Draw earthquake timeline with labels

The follong snippet draws earthquakes observed in the US and Canada from 2000
```{r, fig.show='hold', message=FALSE, warning=FALSE}
    read_eq_clean_data() %>%
        dplyr::filter(YEAR > 2000 & !IS_BC & COUNTRY %in% c("USA", "CANADA")) %>%
        dplyr::mutate(TOTAL_DEATHS = as.numeric(TOTAL_DEATHS),
                      EQ_PRIMARY = as.numeric(EQ_PRIMARY)) %>%
        ggplot2::ggplot(ggplot2::aes(x = DATE,
                                     y = COUNTRY,
                                     colour = TOTAL_DEATHS,
                                     size = EQ_PRIMARY
                                     )) +
        geom_timeline() +
	theme_timeline() +
	labs(size = "Richter scale value", color = "# deaths") +
	geom_timeline_label(ggplot2::aes(label = LOCATION_NAME), n_max = 3)

```

## Visualize the earthquakes on a map with leaflet

The follong snippet map the earthquakes observed in the Mexico from 2000
```{r, fig.show='hold', message=FALSE, warning=FALSE}
    read_eq_clean_data() %>%
        dplyr::filter(YEAR > 2000 & !IS_BC & COUNTRY %in% c("MEXICO")) %>%
        dplyr::mutate(popup_text = eq_create_label(.)) %>%
        eq_map(annot_col = "popup_text")
```
